### 2022-05-10 paper/localpsf.tex
fix typo

In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems of the form
\begin{equation}
\label{eq:Ax_equals_b}
\mathbf{B} \mathbf{u} = \mathbf{f},
\end{equation}
where each of these linear systems is solved using an iterative method. If $\mathbf{B}=\mathbf{A}$, then $\mathbf{A}_H^{-1}$ may be used to precondition this linear system. If $\mathbf{B}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}$. For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}$ is the Hessian of the objective function in the optimization problem, and $\mathbf{f}$ is the negative gradient of the objective function. Both the Hessian and the gradient are evaluated at the current Newton iterate, so $\mathbf{B}$ and $\mathbf{f}$ change as the Newton iterations proceed. The Hessian is typically the sum of a data-misfit term which is available matrix-free, and a regularization term which is available directly as a sparse matrix. The data misfit term in the Hessian is approximated with our method to form $\mathbf{A}_H$, the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian.

Let 
\begin{equation*}
\mathbf{P} \approx \mathbf{B}^{-1}
\end{equation*}
denote the H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method. The iterative method will typically involve, as a sub-procedure, applying $\mathbf{B}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B} \mathbf{x}_i\}_{i=1}^m$. Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, in order to update the preconditioner $P$ for the next linear system. In detail, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively, so that the information learned about $\mathbf{B}$ while solving the current linear system may be summarized by the equation:
\begin{equation*}
\mathbf{B}\mathbf{X} = \mathbf{Y}.
\end{equation*}
For convenience, define 
\begin{equation*}
\mathbf{R} := \mathbf{X}-\mathbf{P}\mathbf{Y}, \qquad \text{and} \qquad \boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}.
\end{equation*}
One of the following three formulas may be used to perform a low rank update of $\mathbf{P}$:
\begin{align}
	\mathbf{P} &\gets \mathbf{P} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
	\mathbf{P} &\gets \mathbf{P} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
	\mathbf{P} &\gets \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T \label{eq:dfp_update}
\end{align}
In all three cases, direct calculation shows that we have $\mathbf{P}\mathbf{Y}=\mathbf{X}$ after the update, which implies $\mathbf{P}^{-1} \mathbf{X} = \mathbf{B}\mathbf{X}$. Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$k$ updates of $\mathbf{P}$, while formula \eqref{eq:dfp_update} performs a rank-$2k$ update. Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}$ is not symmetric. Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}$ is symmetric and indefinite. Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}$ is symmetric positive definite. Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$k$'' update formula. Formula \eqref{eq:dfp_update} is known as the DFP update formula, named after it's inventor Davidson, and polularizers Fletcher and Powell. For more details on these update formulas, we recommend the discussion in Chapter 6 of \cite{NocedalWright99}. In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.

>>>>

In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems of the form
\begin{equation}
\label{eq:Ax_equals_b}
\mathbf{B} \mathbf{u} = \mathbf{f},
\end{equation}
where each of these linear systems is solved using an iterative method. If $\mathbf{B}=\mathbf{A}$, then $\mathbf{A}_H^{-1}$ may be used to precondition this linear system. If $\mathbf{B}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}$. For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}$ is the Hessian of the objective function in the optimization problem, and $\mathbf{f}$ is the negative gradient of the objective function. Both the Hessian and the gradient are evaluated at the current Newton iterate, so $\mathbf{B}$ and $\mathbf{f}$ change as the Newton iterations proceed. The Hessian is typically the sum of a data-misfit term which is available matrix-free, and a regularization term which is available directly as a sparse matrix. The data misfit term in the Hessian is approximated with our method to form $\mathbf{A}_H$, the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian. Let 
\begin{equation*}
\mathbf{P} \approx \mathbf{B}^{-1}
\end{equation*}
denote the H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method. The iterative method will typically involve, as a sub-procedure, applying $\mathbf{B}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B} \mathbf{x}_i\}_{i=1}^m$. Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, in order to update the preconditioner $\mathbf{P}$ for the next linear system. In detail, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively, so that the information learned about $\mathbf{B}$ while solving the current linear system may be summarized by the equation:
\begin{equation*}
\mathbf{B}\mathbf{X} = \mathbf{Y}.
\end{equation*}
For convenience, define 
\begin{equation*}
\mathbf{R} := \mathbf{X}-\mathbf{P}\mathbf{Y}, \qquad \text{and} \qquad \boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}.
\end{equation*}
One of the following three formulas may be used to perform a low rank update of $\mathbf{P}$:
\begin{align}
	\mathbf{P} &\gets \mathbf{P} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
	\mathbf{P} &\gets \mathbf{P} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
	\mathbf{P} &\gets \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T \label{eq:dfp_update}
\end{align}
In all three cases, direct calculation shows that we have $\mathbf{P}\mathbf{Y}=\mathbf{X}$ after the update, which implies $\mathbf{P}^{-1} \mathbf{X} = \mathbf{B}\mathbf{X}$. Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$k$ updates of $\mathbf{P}$, while formula \eqref{eq:dfp_update} performs a rank-$2k$ update. Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}$ is not symmetric. Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}$ is symmetric and indefinite. Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}$ is symmetric positive definite. Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$k$'' update formula. Formula \eqref{eq:dfp_update} is known as the DFP update formula, named after it's inventor Davidson, and polularizers Fletcher and Powell. For more details on these update formulas, we recommend the discussion in Chapter 6 of \cite{NocedalWright99}. In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.
